---
title: "CovidSurg Analysis"
author: "Laura Bravo Merodio"
output: 
  html_document:
    #theme: sandstone
    theme: lumen
    #code_folding: "hide"
    number_sections: TRUE
    toc: true
    toc_float:
        collapsed: true
        smooth_scroll: true
    toc_depth: 3
    fig_caption: yes
    code_folding: show

fontsize: 14pt
---

```{r setup, include=FALSE}

#knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

```


# Introduction

This is a report that summarizes the main results from the research, including plots and tables. It is set so as to mimic the analysis pipeline undergone in the paper *****, where slurm was used and so no reproducibility could be performed. For more information please go to: 

https://globalsurg.org/covidsurg/#Anchor-URL-link

and check our shiny webb app to use our model for COVID-19 patients undergoing surgery. 

https://inflam.shinyapps.io/CovidSugRiskApp/#Anchor-URL-link

***
***

# Set-up and libraries

```{r}

#install.packages("flextable")
library(flextable)
library(knitr)
#install.packages("cowplot")
library(ggplot2)
library(grid)
library(cowplot)
library(dplyr)
#install.packages("webshot")
#install.packages("magick")
library(webshot)
library(magick)
library(officer)
library(gridExtra)
library(magrittr)
library(tibble)
#library(gtsummary); 
library(gt); 
library(forcats);
library(compareGroups)
library(tidyverse)
library(officer)

# load all targets from drake and extras

#loadd(MatrixCorr)
loadd(data2)
loadd(PlotsNames)
loadd(Logitplot)
loadd(TopModels)
loadd(AUCPlots)
loadd(CalibrationPlots)
loadd(hh) 
load("ExternalParams.RData")

```

# Data exploration 

## General information coding

```{r results = 'hide'}

dataA <- readd(data1)

dataPercent <- dataA %>%
  filter(age != "17-19 years")


Info <- data.frame(table(dataPercent$redcap_data_access_group)) %>%
  separate(Var1, into = c("Country","City","Hospital"),sep = "_")

dim(Info) #Hospitals
dim(table(Info$Country)) #Countries

#filter(Info, Country == "es")
table(dataPercent$sex, useNA = "ifany")
table(dataPercent$agegroup, useNA = "ifany")
table(dataPercent$time, useNA = "ifany")
table(dataPercent$urgency4, useNA = "ifany")
table(dataPercent$indication,  useNA = "ifany")
table(dataPercent$spec2, useNA = "ifany")
table(dataPercent$mortality, useNA = "ifany")
table(dataPercent$month, useNA = "ifany")
```

## Preprocessing and table 1

```{r Table, results= 'hide',include=FALSE}

 dataA1Old <- dataA %>%
    dplyr::rename(RespPreop = preop_resp) %>%
    dplyr::rename(Label = mortality) %>%
    dplyr::rename(FinAnas = anaesthesia) %>%
    mutate(
      cardiacrisk = recode(cardiacrisk,
        "0" = "0",
        "1" = "1",
        "2" = "2",
        "3" = "3",
        "4" = "4+",
        "5" = "4+",
        "6" = "4+"
      ),
      asa3 = recode(asa4,
        "Grades 1" = "Grade 1",
        "Grade 2" = "Grade 2",
        "Grade 3" = "Grade 3",
        "Grade 4" = "Grade 4-5",
        "Grade 5" = "Grade 4-5"
      ),
       age = recode(age, '20-29 years' = "<40 years",
                   '30-39 years' = "<40 years")
    ) %>%
    select(-c(
      record_id,
      #redcap_data_access_group,
      # crp,
      starts_with("anaes"),
      starts_with("respsupport"),
      agegroup,
      asa4,
      asa2 # ,
      # month
    )) %>%
    dplyr::rename(anaesthesia = FinAnas) %>%
    # drop_na() %>%
    filter(age != "17-19 years")

  dataA1Old$cardiacrisk <- as.factor(dataA1Old$cardiacrisk)
  
  dataA1Labs <- dataA1Old %>%
    mutate(
      spec2 = recode(spec2,
        "Abdominal" = "Abdominal surgery",
        "Head and neck" = "Head and neck surgery",
        "Limb" = "Orthopaedic surgery",
        "obstetric" = "Obstetrics",
        "other" = "Other surgery",
        "Thorax" = "Cardiothoracic surgery"
      ),
      RespPreop = recode(RespPreop,
        "none" = "None",
        "oxygen" = "Supplemental oxygen",
        "ventilated" = "Ventilation"
      ),
      time = recode(time,
        "Preop" = "Preoperative",
        "Postop" = "Postoperative"
      ),
      indication = recode(indication, "Benign/Obs" = "Benign"),
      grade998 = recode(grade998,
        "major" = "Major",
        "minor" = "Minor"
      ), 
      smoker = recode(smoker, 
                      "No" = "Non-smoker", 
                      "Yes" = "Smoker"), 
      respdisease = recode(respdisease, 
                      "Yes" = "Respiratory disease", 
                      "No" = "No respiratory disease")
    ) %>%
    dplyr::rename(
      Age = age,
      Month = month,
      Sex = sex,
      "Grade of surgery" = grade998,
      Anaesthesia = anaesthesia,
      Speciality = spec2,
      "Haemoglobin (g/dl)" = hb,
      "White cell count (10^9/L)" = wcc,
      "Preoperative respiratory support" = RespPreop,
      "Indication for surgery" = indication,
      "Urgency of surgery" = urgency4,
      "Timing of SARS-CoV-2 diagnosis" = time,
      "Current smoker" = smoker,
      "Respiratory comorbidity" = respdisease,
      "Revised Cardiac Risk Index" = cardiacrisk,
      "ASA grade" = asa3,
      "C-reactive protein" = crp,
    )
  
  data <- dataA1Labs %>%
    filter(Month %in% c("Feb", "Mar", "Apr", "May")) %>%
    select(-Month)

  Validation <- filter(dataA1Labs, Month %in% c("Jun", "July")) %>%
    select(-Month)

#for some reason factors with No and Yes do not work

Total <- list(data, Validation) %>%
  bind_rows(.id = "Type") %>%
  mutate(Type =
  recode(Type, "1" = "Derivation", 
                 "2" = "ValidationTime")
  ) %>%
  mutate_at(c("Current smoker", "Respiratory comorbidity"), as.factor) %>%
  drop_na(Label)
  
res <- compareGroups(Label ~ . - Type, Total, method = 4, chisq.test.perm = TRUE, chisq.test.seed = 132, simplify = TRUE, byrow = FALSE, oddsratio.method ="fisher")
restab1<-createTable(res, hide.no = 'no')  
restab0<-update(restab1, show.n = TRUE, show.ratio = TRUE, show.p.overall = TRUE, show.p.mul = TRUE, show.p.ratio = FALSE)
restab2<-update(restab1, x = update(res, subset = Type == 'ValidationTime'), show.ratio = FALSE, show.p.overall = TRUE, show.p.mul = TRUE, show.p.ratio = FALSE)
restab3<-update(restab1, x = update(res, subset = Type == 'Derivation'), show.ratio = FALSE, show.p.overall = TRUE, show.p.mul = TRUE, show.p.ratio = FALSE)
 p <- cbind("All" = restab0, "Derivation" = restab3, "Validation" = restab2)

export2word(p, file='Tab1_FinalTable.docx')



```

```{r Table1, fig.align="center"}
head(p)
```

## General distribution of categorical variables

```{r DistCat, fig.align="center", fig.width=12, fig.height=8}

Plot1a((data2)[[1]]) + theme_bw()

```

# Feature Selection 

```{r FeatSel1, include = FALSE, results='hide'}

#this grid.arrange generates output!!!


  
  Plot1 <- grid.arrange(
  PlotsNames[[2]] + 
    theme(text = element_text(size=9)), 
  PlotsNames[[4]] + 
    theme(text = element_text(size=10)),
  #Plots[[4]] + 
  #  theme(text = element_text(size=10)),
  #ncol = 3, 
  ncol = 2,
  top = "Variable Importance", 
  #widths = c(0.7, 1,0.8)
  widths = c(0.6, 1)
)



ft_raster <- PlotsNames[[1]] %>% 
  flextable() %>% autofit() %>% add_header_lines("Feature Ranking") %>% 
  as_raster()

gg2 <- ggplot() + 
  theme_void() + 
  annotation_custom(rasterGrob(ft_raster), xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf)

pdf("Fig2_FeatureSelection.pdf",12,7)
cowplot::plot_grid(Plot1, gg2, nrow = 2, ncol = 1, rel_heights = c(4, 2), labels = c('A', 'B'))
dev.off()


```

```{r FeatSel, fig.align="center", fig.width=12, fig.height=14, fig.cap="Figure 2: Feature Selection A. Importance ranking RF and EN. B. Top ranking features "}

cowplot::plot_grid(Plot1, gg2, nrow = 2, ncol = 1, rel_heights = c(4, 2), labels = c('A', 'B'))

```


# Model building

```{r include = F}

load("TrainingTest.RData")
load("FeatureSelect2.RData")

Size <- 12
#Bootstraps carried on

All <- MergeOut %>%
  mutate(Model=
           recode(Model, "dt" = "Decision Trees", 
                  "rf" = "Random Forest", 
                  "logit" = "Logistic Regression")) %>%
  #filter(.metric == "roc_auc") %>%
  ggplot(aes(
    x = Run,
    y = Mean,
    col = Model,
    shape = Type
  )) +
  geom_pointrange(aes(ymin = Mean - sd, ymax = Mean + sd),
                  position = position_dodge(width = 0.8),
                  size = 0.4) +
  facet_grid(Metrics ~ ., scales = "free") +
  scale_y_continuous(
    labels = scales::percent,
    breaks = seq(0, 10, 0.05),
    expand = c(0, 0)
  ) +
  scale_x_continuous(breaks = seq(0, max(MergeOut$Run) + 2, 2), expand = c(0, 0)) +
  labs(title = paste0("Bootstraps n=", Bootstraps)) +
  theme(text = element_text(size = Size)) + theme(
    plot.title = element_text(size = Size, face = "bold"),
    #legend.title = element_text(Size),
    legend.text = element_text(size = Size),
    strip.text = element_text(size = Size),
    panel.spacing = unit(3, "lines")
  ) + theme_bw()


q3 <- MergeOut %>%
  filter(Type == "Test", Metrics == "AUROC") %>%
  select(-c(.metric, Type, ModelFinal)) %>%
  arrange(desc(Mean)) %>%
  mutate_if(is.numeric, ~ round(., 3))

IPP <- inner_join(q3, Variables)

pp <- IPP %>%
  #head(20) %>% #nooooo for the whole table
  #arrange(desc(-Run)) %>%
  mutate(Model=
           recode(Model, "dt" = "Decision Trees", 
                  "rf" = "Random Forest", 
                  "logit" = "Logistic Regression")) %>%
  rename("Standard Deviation" = "sd", 
         #"Number of Variables" = "NumVars", 
         "Mean AUC" = "Mean"
  ) %>%
  select(-Metrics) 

pp$Names <- pp$Names %>%
  str_replace_all(
    c("\"" = "", 
      c("\"" = ""))
  )

l <- Tablize(pp)

pdf("STab2_AllModelsAllRank.pdf", 10,8)
plot(l)
dev.off()


pdf("Fig3A_Model.pdf",18,10)
cowplot::plot_grid(All, NULL, Logitplot, ncol = 1, nrow = 3, rel_heights = c(1.2,0.05, 2), rel_widths  = c(2, 1) )
dev.off()
```

```{r ModelBuild, fig.align="center", fig.width=10, fig.height=8, fig.cap="Figure 3: Model building"}

cowplot::plot_grid(All, NULL, Logitplot, ncol = 1, nrow = 3, rel_heights = c(1.2,0.05, 2), rel_widths  = c(2, 1) )

```

## Model results tables

```{r TablesRes, fig.align="center", fig.width=2, fig.height=1, fig.cap="Table 2"}

library(readr)

read_csv("Top20Models.csv") %>%
                select(-X1) %>%
kable(., digits = 2, format.args = list(scientific = FALSE))


```

## AUCPlots


```{r AUCPlots, fig.align="center", fig.width=7, fig.height=7, fig.cap="Table 2"}

plot(AUCPlots[[3]])

```

## Calibration


```{r Calibration1,include = F}


set.seed(132)

CalibrationParams <- function(ValidationWPred) {
  val_m1 <- val.prob(ValidationWPred$G, as.numeric(as.factor(ValidationWPred$Label)) - 1,
                     pl = FALSE
  ) %>% round(3)
  return(val_m1)
}

CalibrationVals <- rbind(Validation = CalibrationParams(AUCPlots[[2]]), Derivation = CalibrationParams(AUCPlots[[1]]))

CalibrationValsTable <- data.frame(CalibrationVals[,c( 11, 12,13)]) %>%
  add_rownames()

names(CalibrationValsTable)[c(1,2)] <- c("Set",  "Brier")

cc <- Tablize(CalibrationValsTable)

pdf("Fig5C_ParamsCalibration.pdf", 7, 5)
plot(cc)
dev.off()

k <- Plot1Cal(AUCPlots[[2]], " ")

g <- arrangeGrob(k[[1]], k[[2]], respect = TRUE, heights = c(1, 0.25), ncol = 1)

Two <- Plot3Bee(AUCPlots[[2]], "Probability")

ft_raster <- CalibrationValsTable %>% 
  flextable() %>% autofit()  %>% 
  as_raster()

gg2 <- ggplot() + 
  theme_void() + 
  annotation_custom(rasterGrob(ft_raster), xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf)

bottom <- cowplot::plot_grid(Two, gg2, nrow = 2, ncol = 1, rel_heights = c(2, 1), labels = c('B', 'C') )

pdf("Fig5_CalibrationValidation.pdf", 7, 5)
plot_grid(
 g, bottom,
   ncol = 2, align = "h", labels = c('A', '') 
)
dev.off()



```

```{r Calibration, fig.align="center", fig.width=6, fig.height=6, fig.cap="Figure 5: Calibration"}

plot_grid(
 g, bottom,
   ncol = 2, align = "h", labels = c('A', '') 
)

```

# Case fatality and probability outcome

```{r FinalFig1, include = F}

m <- AUCPlots[[4]]

#ChosenRun <- 12  ## get it directly from data. Is it needed?

xx <- data2[[1]] %>%
  select(hh[[ChosenRun]])

vars <- lapply(1:dim(xx)[2], function(x) levels(as.factor(xx[[x]])))

Combination <- expand.grid(vars) %>%
  set_names(names(xx))

CombinationWPred <- EqsFun(Combination, Combination, m)

#max(CombinationWPred$G) # round to 0 
#min(CombinationWPred$G) # round to 0.8
#hist(CombinationWPred$G)

Probs <- round(c(0,0.05, 0.2, 0.5, 0.8), 2)

qqVal <- MortalityTable( Probs, AUCPlots[[2]],hh) 
qqVal %>%
  kable(., digits = 2, format.args = list(scientific = FALSE))

# this one was okay with a 0.55
qqDerv <-  MortalityTable( Probs, AUCPlots[[1]], hh)
qqDerv %>%
  kable(., digits = 2, format.args = list(scientific = FALSE))

write.csv(qqVal, "App-MortalityTableVal.csv")
write.csv(qqDerv, "App-MortalityTableDerv.csv")

priors2 <- data.frame(CombinationWPred) %>%
  mutate(z =cut(G, round( Probs, 2), include.lowest = TRUE)) %>%
  arrange(.,G) 
#had another one with porbabilities instead of linear predictors.


# Case fatality and Linear Predictor

#xxD <- dataA1 %>%
#  select(c(hh[[ChosenRun]]))

DataWPred <- EqsFun(xx, data2[[1]], m)

CaseFatalVal <- DataWPred %>% # have same cuts as before, just want to extract case fatality ratio from here
  mutate(z = cut(G, round( Probs, 2), include.lowest = TRUE)) %>%
  group_by(z) %>%
  mutate(CaseFatality = length(Label[Label == "Dead"]) / (length(Label[Label == "Dead"]) + length(Label[Label == "Alive"]))) %>%
  select(hh[[ChosenRun]], z, CaseFatality) %>% 
  arrange(.,CaseFatality) %>%
  mutate(Risk = paste0(z, "  - ", round(CaseFatality, 3)*100,"%"))
  #mutate(Risk = paste0(round(CaseFatality, 3)*100,"%", z, "  - "))

CaseFatalVal2 <- CaseFatalVal %>%
  select(z, Risk) 

FinalPlot <- inner_join(priors2, CaseFatalVal2) %>%
  unique() 
  
FinalPlot$Risk <- as.factor(FinalPlot$Risk) 

FinalPlot$Risk <-   fct_relevel(FinalPlot$Risk,levels(as.factor(FinalPlot$Risk))[4])

ColoursMort <- ggplot(FinalPlot, aes(x = FinalPlot[[1]], y = FinalPlot[[2]], fill = FinalPlot[[8]])) +
  geom_tile(color = "black", size = 0.5) +
  # geom_text(aes(label=text), size=18) +
  scale_fill_brewer(palette = "Set3") +
  # scale_x_discrete(position = "top") +
  scale_x_discrete("Age") +
  scale_y_discrete("Asa Grade") +
  #scale_fill_discrete(name = "Risk", labels = c("[0,0.2]", "(0.2,0.4]", "(0.4,0.6]", "(0.6,0.8]")) +
  scale_fill_manual(values= c(  "#8FB89D", "#F1D64D", "#E8B49A", "#C57472", "#C94135")) + #"#C94135",
  # values= c( "#b7aeae", "#c3a4a2", "#cf9a96", "#db908a" ,"#e6857f" , "#C57472", "#fe7167")) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1), panel.background = element_blank(), panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), plot.background = element_blank(),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 10, face = "bold"),
    strip.text.x = element_text(size = 10)
  ) +
  facet_grid(FinalPlot[[3]] ~ FinalPlot[[4]], scales = "free" ) + #labeller = label_both
  labs(fill = "Probability Threshold and Mortality")

# do this of whole dataset - not all combinations existing
pdf("Fig6_MortalityTables.pdf", 5,6)
plot(add_header_lines(Tablize(qqVal[-c(3:4),]), values = "Validation", top = TRUE))
plot(add_header_lines(Tablize(qqDerv[-c(3:4),]), values = "Derivation", top = TRUE))
dev.off()

pdf("Fig6_FinalFig.pdf", 14,12)
print(ColoursMort)
dev.off()

```

```{r FinalFigure1, fig.align="center", fig.width=13, fig.height=11, fig.cap="Figure 6: Case fatality ratio"}

ColoursMort 

qqVal[c(1,2,5,6),] %>%
  kable(., digits = 2, format.args = list(scientific = FALSE))

qqDerv[c(1,2,5,6),] %>%
  kable(., digits = 2, format.args = list(scientific = FALSE))

```



